---
title: "SPARCS Data Exploration"
author: "Rose Calixte, PhD, PSTAT"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
---

### Load the required Library

```{r}
library(readr)
require(dplyr)
require(stringr)
require(ggplot2)
require(rpart)
require(chattr)
library(janitor)
library(gtsummary)
require(tidymodels)
require(nnet)
require(glmnet)
require(randomForest)
require(caret)
require(xgboost)
require(ranger)
require(scales)
require(quantreg)
library(purrr)
library(yardstick)
library(vetiver)
library(pins)
library(jsonlite)
library(xgboost)
```

### Loading the data
```{r}

sparcs_2024 <- read_csv("../data/2024/Hospital_Inpatient_Discharges__SPARCS_De-Identified___2024.csv")

sparcs_2023 <- read_csv("../data/2023/Hospital_Inpatient_Discharges__SPARCS_De-Identified___2023.csv")

sparcs_2022 <- read_csv("../data/2022/Hospital_Inpatient_Discharges__SPARCS_De-Identified___2022.csv")

sparcs_2021 <- read_csv("../data/2021/Hospital_Inpatient_Discharges__SPARCS_De-Identified___2021.csv")

attach(sparcs_2021)
attach(sparcs_2022)
attach(sparcs_2023)
attach(sparcs_2024)
```

### Clean, rename, a nd merge the master files

```{r}
# ==============================================================================
# MASTER PIPELINE: INDIVIDUAL CLEANING (2021-2024) -> MERGE
# ==============================================================================

# --- 2021 Data Cleaning ---
clean_2021 <- clean_names(sparcs_2021) %>%
  mutate(
    # 1. Clean and convert birth_weight
    birth_weight = as.numeric(str_trim(as.character(birth_weight))),
    
    # 2. Clean and convert length_of_stay
    los_clean    = str_trim(str_remove(as.character(length_of_stay), "\\+")),
    length_of_stay = as.numeric(los_clean),
    
    # 3. Clean and convert total_charges (strips out commas, dollar signs, and spaces)
    charges_clean = str_remove_all(as.character(total_charges), "[\\$, ]"),
    total_charges = as.numeric(charges_clean),
    
    # 4. Clean and convert total_costs (strips out commas, dollar signs, and spaces)
    costs_clean   = str_remove_all(as.character(total_costs), "[\\$, ]"),
    total_costs   = as.numeric(costs_clean)
  ) %>%
  # Drop all the temporary intermediate calculation columns
  select(-los_clean, -charges_clean, -costs_clean)

# --- 2022 Data Cleaning ---
clean_2022 <- clean_names(sparcs_2022) %>%
  mutate(
    birth_weight = as.numeric(str_trim(as.character(birth_weight)))
  ) 

# --- 2023 Data Cleaning ---
clean_2023 <- clean_names(sparcs_2023) %>%
  mutate(
    birth_weight = as.numeric(str_trim(as.character(birth_weight)))
  ) 

# --- 2024 Data Cleaning (Just Length of Stay + Name Fixes) ---
clean_2024 <- clean_names(sparcs_2024) %>%
  # 1. Fix the 2024 column name shifts first
  rename(
    hospital_service_area = health_service_area,
    zip_code_3_digits     = zip_code
  ) %>%
  # 2. Process JUST length_of_stay for 2024
  mutate(
    los_clean    = str_trim(str_remove(as.character(length_of_stay), "\\+")),
    length_of_stay = as.numeric(los_clean)
  ) %>%
  select(-los_clean)

# ==============================================================================
# FINAL MASTER MERGE
# ==============================================================================

# Stacks all four years seamlessly since columns and data types now line up perfectly
sparcs_master <- bind_rows(clean_2021, clean_2022, clean_2023, clean_2024)

# ==============================================================================
# DATA TYPE VERIFICATION CHECKS
# ==============================================================================
print("Final Column Classes in Master Dataset:")
print(paste("Length of Stay:", class(sparcs_master$length_of_stay)))
print(paste("Birth Weight:", class(sparcs_master$birth_weight)))
print(paste("Total Charges:", class(sparcs_master$total_charges)))
print(paste("Total Costs:", class(sparcs_master$total_costs)))

table(sparcs_master$payment_typology_1)
```

### Filter to NYC Area, Vaginal Delivery, Private Insurance or BCBS

```{r}
# ==============================================================================
# COHORT SELECTION: NYC, PRIVATE/BCBS INSURANCE, VAGINAL DELIVERY
# ==============================================================================

sparcs_filtered <- sparcs_master %>%
  filter(
    # 1. Restrict geographically to New York City
    hospital_service_area == "New York City",
    
    # 2. Restrict to Primary Insurance that is either BCBS or Private Health Insurance
    payment_typology_1 %in% c("Blue Cross/Blue Shield", "Private Health Insurance"),
    
    # 3. Isolate vaginal deliveries using the APR-DRG description
    # (Using grepl for a case-insensitive search handles any minor label shifts)
    grepl("Vaginal Delivery", apr_drg_description, ignore.case = TRUE)
  )

sparcs_filtered <- sparcs_filtered %>%
  filter(
    # Converts the column to uppercase and looks for an exact or clean match
    toupper(apr_drg_description) == "VAGINAL DELIVERY"
  )
# ==============================================================================
# CLEAN AND UNIFY AGE GROUPS
# ==============================================================================

sparcs_filtered <- sparcs_filtered %>%
  mutate(
    # 1. Convert " to " into "-" and strip out extra spaces
    age_group = str_replace_all(age_group, " to ", "-"),
    age_group = str_trim(age_group)
  ) %>%
  # 2. Filter out the clinical anomalies (ages 50-69) for a live birth cohort
  filter(age_group != "50-69")
sparcs_filtered <- sparcs_filtered %>%
  filter(
    # Keep only the adult age brackets
    age_group != "0-17"
  )

sparcs_filtered <- sparcs_filtered %>%
  filter(
    # Keep only the rows matching your three target categories
    type_of_admission %in% c("Elective", "Emergency", "Urgent")
  )

sparcs_filtered <- sparcs_filtered %>%
  filter(
    # Kepp only those with minor risk of mortality
    toupper(apr_risk_of_mortality) =="MINOR"
  )
sparcs_filtered <- sparcs_filtered %>%
  filter(
    # Filter out those with extreme severity of illness
    toupper(apr_severity_of_illness_description) !="EXTREME"
  )
# ==============================================================================
# VERIFICATION CHECK
# ==============================================================================

print("Remaining types of admission:")
print(table(sparcs_filtered$type_of_admission))

print("Cleaned and unified age groups:")
print(table(sparcs_filtered$age_group))

print("Breakdown of Primary Payer in final cohort:")
table(sparcs_filtered$payment_typology_1)

print("Unique Delivery Types remaining:")
table(sparcs_filtered$apr_drg_description)
```

### Creating the Ratio indicator variable

```{r}
# ==============================================================================
# ADD FINANCIAL METRICS: COST-TO-CHARGE PERCENTAGE & 50% INDICATOR
# ==============================================================================

sparcs_filtered <- sparcs_filtered %>%
  mutate(
    # 1. Calculate the cost as a percentage of the total charge
    cost_to_charge_pct = (total_costs / total_charges) * 100,
    
    # 2. Create an indicator variable (1 = 50% or more, 0 = Less than 50%)
    cost_gte_50_pct = if_else(cost_to_charge_pct >= 50, 1, 0),
    
    # 3. Create a Cost to charge differential
    cost_to_charge_diff = total_costs-total_charges
  )

# ==============================================================================
# VERIFICATION CHECKS
# ==============================================================================

print("Summary statistics for the new percentage variable:")
print(summary(sparcs_filtered$cost_to_charge_pct))

print("Breakdown of the 50% or more indicator flag:")
print(table(sparcs_filtered$cost_gte_50_pct, useNA = "ifany"))
```

### Categorize the hospitals into private, public, and community

```{r}

# ==============================================================================
# CREATE THREE-TIER HOSPITAL TYPE VARIABLE
# ==============================================================================

sparcs_filtered <- sparcs_filtered %>%
  mutate(
    # Create an uppercase placeholder for matching to eliminate duplicate case rows
    fac_upper = toupper(facility_name),
    
    # Define our three tiers based on operational structure
    hospital_tier = case_when(
      
      # 1. PUBLIC HOSPITALS (NYC H+H & State SUNY)
      grepl("BELLEVUE|CONEY ISLAND|SOUTH BROOKLYN HEALTH|ELMHURST|HARLEM|JACOBI|KINGS COUNTY|LINCOLN|METROPOLITAN|NORTH CENTRAL BRONX|QUEENS HOSPITAL|WOODHULL|UNIVERSITY HOSPITAL OF BROOKLYN", fac_upper) 
      ~ "Public",
      
      # 2. PRIVATE SYSTEMS & ACADEMIC MEDICAL CENTERS
      grepl("PRESBYTERIAN|NYU LANGONE|MOUNT SINAI|LENOX HILL|LONG ISLAND JEWISH|STATEN ISLAND UNIVERSITY|MONTEFIORE|MAIMONIDES", fac_upper) 
      ~ "Private System",
      
      # 3. COMMUNITY / SAFETY-NET VOLUNTARY HOSPITALS (Everything else remaining)
      TRUE ~ "Community"
    ),
    
    # Convert to factor for your downstream logistic/GLM modeling
    hospital_tier = factor(hospital_tier, levels = c("Public", "Community", "Private System"))
  ) %>%
  # Drop our temporary uppercase matching column
  select(-fac_upper)

# ==============================================================================
# VERIFICATION CHECK
# ==============================================================================
print("Count of records in each hospital tier:")
print(table(sparcs_filtered$hospital_tier))

sparcs_filtered <- sparcs_filtered %>%
  mutate(
    # If the county is "Manhattan", overwrite it with "New York"
    hospital_county = if_else(hospital_county == "Manhattan", "New York", hospital_county)
  )

# ==============================================================================
# VERIFICATION CHECK
# ==============================================================================

print("Cleaned and collapsed NYC hospital counties:")
print(table(sparcs_filtered$hospital_county))

write_csv(sparcs_filtered, "sparcs_filtered.csv")
```

### Summarize the data

```{r}
sparcs_filtered %>%
  select(discharge_year, age_group, race, ethnicity, type_of_admission, patient_disposition, hospital_county, hospital_tier,
         apr_severity_of_illness_code, apr_severity_of_illness_description,
         apr_risk_of_mortality, apr_medical_surgical_description,
         payment_typology_1, apr_drg_description, length_of_stay, 
         total_charges, total_costs, cost_to_charge_diff, cost_to_charge_pct, cost_gte_50_pct) %>%
  tbl_summary(
    missing = "ifany", # Displays row for missing values only if they exist
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",      # Mean and standard deviation for numeric
      all_categorical() ~ "{n} ({p}%)"         # Counts and percentages for categories
    )
  ) %>%
  bold_labels() 
```

### Compare the characteristics across year

```{r}
sparcs_filtered %>%
  select(
    discharge_year, # <-- Added the grouping variable here
    age_group, race, ethnicity, type_of_admission, hospital_county, hospital_tier,
    apr_severity_of_illness_code, apr_severity_of_illness_description,payment_typology_1, length_of_stay, 
    total_charges, total_costs, cost_to_charge_diff, cost_to_charge_pct, cost_gte_50_pct
  ) %>%
  tbl_summary(
    by = discharge_year, # <-- Groups data into side-by-side columns by year
    missing = "ifany", 
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",  
      all_categorical() ~ "{n} ({p}%)"     
    )
  ) %>%
  bold_labels() %>%
  add_p() # Optional: Adds statistical tests across years (e.g., ANOVA, Chi-Square)
```

### Visualizing the Quantitative Variables as Histograms

```{r}
# --- Histogram 1: Total Charges ---
ggplot(sparcs_filtered, aes(x = total_charges)) +
  geom_histogram(fill = "steelblue", color = "white", bins = 50) +
  scale_x_continuous(labels = label_dollar()) + # Formats numbers as $50,000 etc.
  labs(
    title = "Distribution of Total Charges for Vaginal Deliveries",
    subtitle = "Privately Insured Cohort in NYC",
    x = "Total Charges ($)",
    y = "Count of Patients"
  ) +
  theme_minimal(base_size = 14)

# --- Histogram 2: Total Costs ---
ggplot(sparcs_filtered, aes(x = total_costs)) +
  geom_histogram(fill = "darkgreen", color = "white", bins = 50) +
  scale_x_continuous(labels = label_dollar()) +
  labs(
    title = "Distribution of Total Costs for Vaginal Deliveries",
    subtitle = "Privately Insured Cohort in NYC",
    x = "Total Costs ($)",
    y = "Count of Patients"
  ) +
  theme_minimal(base_size = 14)

# --- Histogram 3: Length of Stay ---
ggplot(sparcs_filtered, aes(x = length_of_stay)) +
  geom_histogram(fill = "darkorange", color = "white", binwidth = 1) + # 1-day bins
  scale_x_continuous(breaks = seq(0, 10, by = 2)) + # Shows ticks for each day
  labs(
    title = "Distribution of Length of Stay for Vaginal Deliveries",
    subtitle = "Privately Insured Cohort in NYC",
    x = "Length of Stay (Days)",
    y = "Count of Patients"
  ) +
  theme_minimal(base_size = 14)

# ==============================================================================
# HISTOGRAM FOR COST-TO-CHARGE DIFFERENCE
# ==============================================================================

ggplot(sparcs_filtered, aes(x = cost_to_charge_diff)) +
  # Create the histogram bars
  geom_histogram(fill = "purple", color = "white", bins = 50) +
  
  # Add a distinct vertical line at $0 to mark where cost equals charge
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkred", size = 1) +
  
  # Format the x-axis numbers as clean currency (e.g., -$10,000)
  scale_x_continuous(labels = label_dollar()) +
  
  # Titles and metadata
  labs(
    title = "Distribution of Net Financial Difference (Cost - Charge)",
    subtitle = "Vaginal Deliveries in NYC | Dashed line indicates Cost = Charge",
    x = "Financial Difference (Total Costs - Total Charges)",
    y = "Count of Patients"
  ) +
  
  # Polished styling
  theme_minimal(base_size = 14)
```

### Preparing the data for analysis

```{r}
# 1. Format data for classification
modeling_data <- sparcs_filtered %>%
  mutate(
    # Caret requires factors for classification, and valid R variable names for labels
    cost_gte_50_pct = factor(if_else(cost_gte_50_pct == 1, "GTE_50", "LESST_50"))
  ) %>%
  # Select only the target and the specific predictor/control variables you want
  select(discharge_year, age_group,race,ethnicity,length_of_stay,apr_severity_of_illness_description,payment_typology_1,type_of_admission,hospital_county, hospital_tier, total_charges,total_costs, cost_to_charge_pct, cost_gte_50_pct, cost_to_charge_diff)
```

### Partition the Data

```{r}
set.seed(4568)

split <- modeling_data$cost_gte_50_pct %>% 
  createDataPartition(p = 0.7, list = FALSE)

train_data <- modeling_data[split, ]
test_data <- modeling_data[-split, ]
```

### create a test training a model (logistic regression)

```{r}
# Configure cross-validation and request probability outputs
cv_control <- trainControl(
  method = "cv",
  number = 10,
  classProbs = TRUE,                  # Necessary to calculate ROC curve downstream
  summaryFunction = twoClassSummary  # Evaluates using ROC/AUC instead of raw Accuracy
)

# Fit the logistic regression model
logistic_model <- train(
  cost_gte_50_pct ~ discharge_year+age_group+race+ethnicity+length_of_stay+apr_severity_of_illness_description+
    payment_typology_1+type_of_admission+hospital_county+ hospital_tier, 
  data = train_data, 
  method = "glm",
  family = "binomial",                # Tells GLM to run logistic link function
  trControl = cv_control,
  metric = "ROC"                      # Optimizes the model based on Area Under the ROC Curve
)

# View model overview and CV results
print(logistic_model)
```

### Testing the model

```{r}
# 1. Predict class labels on test data
test_predictions <- predict(logistic_model, newdata = test_data)

# 2. Generate Confusion Matrix (Shows Accuracy, Sensitivity, Specificity)
confusionMatrix(test_predictions, test_data$cost_gte_50_pct)

summary(logistic_model$finalModel)

```

### Creating an Ensemble of models for quantitative variables

```{r}
# Define Candidate Models

# 1. Linear Regression
lm_spec <- linear_reg() %>%
  set_engine("lm")

# 3. Ridge Regression
ridge_spec <- linear_reg(mixture = 0, penalty = 0.1) %>%
  set_engine("glmnet")

# 4. Regression Tree
tree_spec <- decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("regression")

# 5. Random Forest
rf_spec <- rand_forest(trees = 100) %>%
  set_engine("ranger") %>%
  set_mode("regression")

# 6. XGBoost (Fixed Typos)
xgb_spec <- boost_tree(trees = 100) %>%  # Fixed function name and argument plural 'trees'
  set_engine("xgboost") %>%              # Fixed engine name
  set_mode("regression")

models = list(
  "Linear Regression" = lm_spec,
  "Ridge Regression" = ridge_spec,
  "Regression Tree" = tree_spec,
  "Random Forest" = rf_spec,
  "XGBoost" = xgb_spec
)
```

### Helper: Save champion models as JSON

```{r}
# ==============================================================================
# HELPER: SAVE CHAMPION MODEL AS JSON (alongside the vetiver pin)
# ==============================================================================
# Writes two files per champion into ml/models:
#   <model_name>.json          - full model serialized with jsonlite::serializeJSON(),
#                                 reload with jsonlite::unserializeJSON()
#   <model_name>_metadata.json - human-readable champion name, formula, and metrics

json_models_dir <- "../models"
if (!dir.exists(json_models_dir)) dir.create(json_models_dir, recursive = TRUE)

save_champion_json <- function(model_name, fitted_workflow, formula, results_table) {
  # 1. Extract the actual fitted xgboost model from the tidymodels workflow
  xgb_model <- extract_fit_parsnip(fitted_workflow)$fit
  
  # 2. Save the XGBoost model natively as a JSON file
  # Note: The extension MUST be .json
  model_path <- file.path(json_models_dir, paste0(model_name, ".json"))
  # xgb.save(xgb_model, model_path, fname_format = "json")
  xgb.save(xgb_model, model_path)
  
  # write(
  #  jsonlite::serializeJSON(fitted_workflow),
  #  file.path(json_models_dir, paste0(model_name, ".json"))
  #)

  metadata <- list(
    model_name = model_name,
    champion   = as.character(results_table$Model[1]),
    formula    = deparse(formula),
    metrics    = results_table,
    saved_at   = as.character(Sys.time())
  )
  write_json(
    metadata,
    file.path(json_models_dir, paste0(model_name, "_metadata.json")),
    auto_unbox = TRUE,
    pretty = TRUE
  )
}
```

### Train, test, and rank the Models for Total Charges

```{r}
# Create an empty list container to keep track of the physical model objects
fitted_charges_models <- list()

results_charges = map_dfr(names(models), function(name) {
  fitted_model = models[[name]] %>%
    fit(total_charges ~ discharge_year+ age_group + race + ethnicity + length_of_stay + apr_severity_of_illness_description +
        payment_typology_1 + type_of_admission + hospital_county + hospital_tier, data = train_data)
  
  # Save the trained model object to our external list before moving to the next iteration
  fitted_charges_models[[name]] <<- fitted_model
  
  preds = predict(fitted_model, new_data = test_data) %>%
    bind_cols(test_data)
  
  rmse_val = mse(preds, truth = total_charges, estimate = .pred) %>% pull(.estimate)
  rsq_value = rsq(preds, truth = total_charges, estimate = .pred) %>% pull(.estimate)
  
  tibble(Model = name, RMSE = rmse_val, Rsq = rsq_value)
})

# ==============================================================================
# 0. INITIALIZE THE STORAGE BOARD (Fixes the 'not found' error)
# ==============================================================================
# This creates or connects to the local folder where your models live
champion_board <- board_folder("saved_champion_models", versioned = TRUE)


# ==============================================================================
# 1. IDENTIFY THE WINNING MODEL BLUEPRINT
# ==============================================================================
# Ensure the results dataframe is cleanly sorted by lowest RMSE
results_charges <- results_charges %>% arrange(RMSE)
results_charges

# Pull the string name of the winner (e.g., "XGBoost")
best_charges_name <- results_charges$Model[1] 

# Pull the fresh, un-fitted specification straight from your original 'models' list
best_charges_spec <- models[[best_charges_name]]


# ==============================================================================
# 2. CREATE A CLEAN TIDYMODELS WORKFLOW & FIT IT
# ==============================================================================
# Creating an explicit workflow container ensures vetiver gets the exact data
# types and formula structures it needs to build its internal description.
champion_workflow <- workflow() %>%
  add_model(best_charges_spec) %>%
  add_formula(
    total_charges ~ discharge_year+age_group + race + ethnicity + length_of_stay + 
                    apr_severity_of_illness_description + payment_typology_1 + 
                    type_of_admission + hospital_county + hospital_tier
  )

# Explicitly train this specific workflow on your training data
champion_charges_fit <- fit(champion_workflow, data = train_data)


# ==============================================================================
# 3. PASS TO VETIVER & WRITE PIN
# ==============================================================================
# Pass the fully trained workflow object directly into the deployment constructor
v_charges <- vetiver_model(
  model = champion_charges_fit, 
  model_name = "delivery_charges_regression"
)

# Write to your local dashboard board directory
champion_board %>% vetiver_pin_write(v_charges)

# Also store the champion model as JSON under ml/models
save_champion_json(
  model_name = "delivery_charges_regression",
  fitted_workflow = champion_charges_fit,
  formula = total_charges ~ discharge_year + age_group + race + ethnicity + length_of_stay +
                    apr_severity_of_illness_description + payment_typology_1 +
                    type_of_admission + hospital_county + hospital_tier,
  results_table = results_charges
)
```

### Train, test, and rank the Models for Total Costs

```{r}
fitted_costs_models <- list()

results_costs = map_dfr(names(models), function(name) {
  fitted_model = models[[name]] %>%
    fit(total_costs ~ discharge_year+age_group + race + ethnicity + length_of_stay + apr_severity_of_illness_description +
        payment_typology_1 + type_of_admission + hospital_county + hospital_tier, data = train_data)
  
  fitted_costs_models[[name]] <<- fitted_model
  
  preds = predict(fitted_model, new_data = test_data) %>%
    bind_cols(test_data)
  
  # FIX: Evaluated against total_costs instead of total_charges
  rmse_val = mse(preds, truth = total_costs, estimate = .pred) %>% pull(.estimate)
  rsq_value = rsq(preds, truth = total_costs, estimate = .pred) %>% pull(.estimate)
  
  tibble(Model = name, RMSE = rmse_val, Rsq = rsq_value)
})

# ==============================================================================
# EXTRACT AND STORE CHAMPION MODEL FOR TOTAL COSTS
# ==============================================================================

# 0. Connect to the storage board
champion_board <- board_folder("saved_champion_models", versioned = TRUE)

# 1. Identify the winning model blueprint from your results dataframe
results_costs <- results_costs %>% arrange(RMSE)
results_costs
best_costs_name <- results_costs$Model[1] 
best_costs_spec <- models[[best_costs_name]]

# 2. Create a clean workflow container and fit it
workflow_costs <- workflow() %>%
  add_model(best_costs_spec) %>%
  add_formula(
    total_costs ~ discharge_year+age_group + race + ethnicity + length_of_stay + 
                  apr_severity_of_illness_description + payment_typology_1 + 
                  type_of_admission + hospital_county + hospital_tier
  )

champion_costs_fit <- fit(workflow_costs, data = train_data)

# 3. Create vetiver model and pin it to the board
v_costs <- vetiver_model(
  model = champion_costs_fit, 
  model_name = "delivery_costs_regression"
)

champion_board %>% vetiver_pin_write(v_costs)

# Also store the champion model as JSON under ml/models
save_champion_json(
  model_name = "delivery_costs_regression",
  fitted_workflow = champion_costs_fit,
  formula = total_costs ~ discharge_year + age_group + race + ethnicity + length_of_stay +
                  apr_severity_of_illness_description + payment_typology_1 +
                  type_of_admission + hospital_county + hospital_tier,
  results_table = results_costs
)
```


### Train, test, and rank the Models for Costs to Charge Differential

```{r}
fitted_diff_models <- list()

results_diff = map_dfr(names(models), function(name) {
  fitted_model = models[[name]] %>%
    fit(cost_to_charge_diff ~ discharge_year+age_group + race + ethnicity + length_of_stay + apr_severity_of_illness_description +
        payment_typology_1 + type_of_admission + hospital_county + hospital_tier, data = train_data)
  
  fitted_diff_models[[name]] <<- fitted_model
  
  preds = predict(fitted_model, new_data = test_data) %>%
    bind_cols(test_data)
  
  # FIX: Evaluated against cost_to_charge_diff instead of total_charges
  rmse_val = mse(preds, truth = cost_to_charge_diff, estimate = .pred) %>% pull(.estimate)
  rsq_value = rsq(preds, truth = cost_to_charge_diff, estimate = .pred) %>% pull(.estimate)
  
  tibble(Model = name, RMSE = rmse_val, Rsq = rsq_value)
})

# ==============================================================================
# EXTRACT AND STORE CHAMPION MODEL FOR NET DIFFERENCE
# ==============================================================================

# 0. Connect to the storage board
champion_board <- board_folder("saved_champion_models", versioned = TRUE)

# 1. Identify the winning model blueprint from your results dataframe
results_diff <- results_diff %>% arrange(RMSE)
results_diff
best_diff_name <- results_diff$Model[1] 
best_diff_spec <- models[[best_diff_name]]

# 2. Create a clean workflow container and fit it
workflow_diff <- workflow() %>%
  add_model(best_diff_spec) %>%
  add_formula(
    cost_to_charge_diff ~ discharge_year+age_group + race + ethnicity + length_of_stay + 
                          apr_severity_of_illness_description + payment_typology_1 + 
                          type_of_admission + hospital_county + hospital_tier
  )

champion_diff_fit <- fit(workflow_diff, data = train_data)

# 3. Create vetiver model and pin it to the board
v_diff <- vetiver_model(
  model = champion_diff_fit, 
  model_name = "delivery_margin_diff_regression"
)

champion_board %>% vetiver_pin_write(v_diff)

# Also store the champion model as JSON under ml/models
save_champion_json(
  model_name = "delivery_margin_diff_regression",
  fitted_workflow = champion_diff_fit,
  formula = cost_to_charge_diff ~ discharge_year + age_group + race + ethnicity + length_of_stay +
                          apr_severity_of_illness_description + payment_typology_1 +
                          type_of_admission + hospital_county + hospital_tier,
  results_table = results_diff
)
```

### Binary Outcome

```{r}

# 1. Logistic Regression (Replaced linear_reg)
logistic_spec <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

# 2. Classification Tree (Changed mode to classification)
tree_spec <- decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("classification")

# 3. Random Forest (Changed mode to classification)
rf_spec <- rand_forest(trees = 100) %>%
  set_engine("ranger") %>%
  set_mode("classification")

# 4. XGBoost (Changed mode to classification)
xgb_spec <- boost_tree(trees = 100) %>%  
  set_engine("xgboost") %>%              
  set_mode("classification")

# ==============================================================================
# UPDATED CANDIDATE LIST
# ==============================================================================
models = list(
  "Logistic Regression"       = logistic_spec,
  "Classification Tree"       = tree_spec,
  "Random Forest"             = rf_spec,
  "XGBoost"                   = xgb_spec
)
```

### Train, Test, and Rank the models for payment ≥ 50%

```{r}

fitted_binary_models <- list()

results_binary = map_dfr(names(models), function(name) {
  fitted_model = models[[name]] %>%
    fit(cost_gte_50_pct ~ discharge_year+age_group + race + ethnicity + length_of_stay + 
        apr_severity_of_illness_description + payment_typology_1 + 
        type_of_admission + hospital_county + hospital_tier, data = train_data)
  
  fitted_binary_models[[name]] <<- fitted_model
  
  hard_preds <- predict(fitted_model, new_data = test_data)
  prob_preds <- predict(fitted_model, new_data = test_data, type = "prob")
  
  preds <- test_data %>% bind_cols(hard_preds, prob_preds)
  
  accuracy_val <- accuracy(preds, truth = cost_gte_50_pct, estimate = .pred_class) %>% pull(.estimate)
  
  auc_val <- suppressWarnings(
    roc_auc(preds, truth = cost_gte_50_pct, .pred_GTE_50) %>% pull(.estimate)
  )
  
  tibble(Model = name, Accuracy = accuracy_val, ROC_AUC = auc_val)
})

# ==============================================================================
# EXTRACT AND STORE CHAMPION MODEL FOR BINARY CLASSIFICATION (≥ 50%)
# ==============================================================================

# 0. Connect to the storage board
champion_board <- board_folder("saved_champion_models", versioned = TRUE)

# 1. Identify the winning classification model blueprint from your results dataframe
results_binary <- results_binary %>% arrange(desc(ROC_AUC))
results_binary
best_binary_name <- results_binary$Model[1] 
best_binary_spec <- models[[best_binary_name]]

# 2. Create a clean classification workflow container and fit it
workflow_binary <- workflow() %>%
  add_model(best_binary_spec) %>%
  add_formula(
    cost_gte_50_pct ~ discharge_year+age_group + race + ethnicity + length_of_stay + 
                      apr_severity_of_illness_description + payment_typology_1 + 
                      type_of_admission + hospital_county + hospital_tier
  )

champion_binary_fit <- fit(workflow_binary, data = train_data)

# 3. Create vetiver model and pin it to the board
v_binary <- vetiver_model(
  model = champion_binary_fit, 
  model_name = "delivery_recovery_threshold_classification"
)

champion_board %>% vetiver_pin_write(v_binary)

# Also store the champion model as JSON under ml/models
save_champion_json(
  model_name = "delivery_recovery_threshold_classification",
  fitted_workflow = champion_binary_fit,
  formula = cost_gte_50_pct ~ discharge_year + age_group + race + ethnicity + length_of_stay +
                      apr_severity_of_illness_description + payment_typology_1 +
                      type_of_admission + hospital_county + hospital_tier,
  results_table = results_binary
)
```


